---
title: <center><font size=30; font-weight=775>COVID-19 Report</font></center>
author: <center>Alejandro Álvarez</center>
date: <center>`r Sys.setlocale('LC_TIME','English'); format(Sys.time(), '%B %d, %Y')`</center>
output:
   html_document:
      toc: true
      toc_depth: 3
      toc_float:
         collapse: true
         smooth_scroll: true
      includes:
         in_header: header.html
         after_body: footer.html
      css: style.css
pagetitle: 'COVID-19 Report'
---

Sources:

* [Our World in Data. Data collected, aggregated, and documented by Diana Beltekian, Daniel Gavrilov, Charlie Giattino, Joe Hasell, Bobbie Macdonald, Edouard Mathieu, Esteban Ortiz-Ospina, Hannah Ritchie, Max Roser](https://github.com/owid/covid-19-data/tree/master/public/data). Accessed: `r Sys.setlocale('LC_TIME','English'); format(Sys.time(), '%B %d, %Y')`.

* [Google LLC 'Google COVID-19 Community Mobility Reports'](https://www.google.com/covid19/mobility/). Accessed: `r Sys.setlocale('LC_TIME','English'); format(Sys.time(), '%B %d, %Y')`.

```{r setup, include=FALSE}

##### R Markdown global options #####

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, include = TRUE, fig.align = 'center', out.width = '100%')

##### libraries #####

library(tidyverse)
library(plotly)
library(extrafont)
library(mgcv)
loadfonts(device = 'win') # The first time you install this package, you have to execute font_import()

##### data import #####

mobility_raw_data <- read_csv('inputs/global_mobility_report.csv')

covid_data <- read_csv('inputs/owid-covid-data.csv')

##### Colors #####

colors_pal <- c('#004170','#098d9c','#cc003a','#b2d600','#ffcc00','#00c954','#00bf9f','#00a9bf','#6f00bf','#b02c88')

colors_continent <- sample(colors_pal,6,replace = FALSE)

##### Dummy function #####

do_nothing <- function(df) { return(df) }

```

# COVID-19 Cases

[Text]

```{r data_frames}

covid1 <- covid_data %>%
   arrange(location,date) %>% 
   group_by(location) %>%
   mutate(max_cases = max(total_cases,na.rm = TRUE),
          max_deaths = max(total_deaths,na.rm = TRUE),
          max_cases_per_million = max(total_cases_per_million,na.rm = TRUE),
          max_deaths_per_million = max(total_deaths_per_million,na.rm = TRUE)) %>%
   filter(location != 'World') %>% 
   do_nothing()

```

## Total Accumulated Cases by Country

[Text]

[Text]

```{r total_cases_by_country_plot, fig.height=8}

covid1 %>% 
   filter( (max(total_cases_per_million,na.rm = TRUE) > 1000 &
               max_cases > 40000 &
               population > 5000000 &
               total_cases >= 100) |
              location == 'China' ) %>%
   ggplot(aes(x = date, y = total_cases/1000, color = continent)) +
   geom_line(size = 2) +
   facet_wrap(~fct_reorder(location,max_cases, .desc = TRUE), scales = 'free_y', nrow = 7, ncol = 4) +
   scale_y_continuous(labels = scales::comma_format(accuracy = 1)) +
   scale_color_manual(values = colors_continent) +
   theme_minimal() +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill = '#fafafa', color = 'white'),
         strip.background = element_rect(fill = '#fafafa', color = 'white'),
         strip.text = element_text(family = 'Century Gothic', color = '#053347'),
         legend.position = 'bottom',
         legend.title = element_blank(),
         axis.title.x = element_blank(),
         axis.text = element_text(family = 'Century Gothic', color = '#053347'),
         text = element_text(family = 'Century Gothic', color = '#053347'),
         plot.title = element_text(hjust = 0.5, face = 'bold'),
         plot.caption = element_text(hjust = 0),
         NULL) +
   ylab('Total cases (\'000)\n') +
   labs(title = 'Total cases\' evolution since the 100th case, by country\n',
        caption = '* Only countries with more than 40,000 cases, more than 1,000 cases per million habitants,\nand with a total population greater than 5 million people are shown (except for China).\n\nData source: Our World in Data.')

```

<br><br>

<center>**Statistics by Country**</center>

<br>

```{r total_cases_by_country_table}

covid1 %>% 
   filter(date == last(date)) %>% 
   select(date, location, total_cases, total_cases_per_million, total_deaths, total_deaths_per_million) %>% 
   arrange(-total_cases) %>% 
   rename_all(funs(c('Date', 'Country', 'Total Cases', 'Total Cases per Million', 'Total Deaths', 'Total Deaths per Million'))) %>% 
   DT::datatable(., rownames = FALSE,
                 filter = 'top',
                 width = 825,
                 options = list(pageLength = 20,
                                scrollX = TRUE,
                                searching = TRUE,
                                lengthMenu = c(10,20,30,40,50),
                                columnDefs = list(list(className = 'dt-center', targets = 0:5)))) %>% 
   DT::formatRound(table = ., columns = c('Total Deaths per Million'),digits = 2) %>% 
   DT::formatCurrency(table = ., columns = c('Total Cases', 'Total Cases per Million', 'Total Deaths'),currency = '', digits = 0) %>%
   DT::formatStyle(0:5,color = '#053347')

```

## New Cases by Country

[Text]

[Text]

```{r new_cases_by_country_p, fig.height=9}

p <- covid1 %>% 
   filter( (max(total_cases_per_million,na.rm = TRUE) > 1000 &
               max_cases > 40000 &
               population > 5000000 &
               total_cases >= 100) |
              location == 'China' ) %>%
   mutate('New cases (mov. av.)' = round(zoo::rollmean(x = new_cases,k = 7, fill = NA, align = 'left'),digits = 0)) %>%
   rename('Date' = date) %>% 
   ggplot(aes(x = Date, y = `New cases (mov. av.)`, color = continent)) +
   geom_line(size = 0.5) +
   facet_wrap(~fct_reorder(location,max_cases, .desc = TRUE), scales = 'free_y', nrow = 7, ncol = 4) +
   scale_y_continuous(labels = scales::comma_format(accuracy = 1)) +
   scale_color_manual(values = colors_continent) +
   theme_minimal() +
   theme(panel.grid = element_blank(),
         panel.background = element_rect(fill = '#fafafa', color = 'white'),
         strip.background = element_rect(fill = '#fafafa', color = 'white'),
         legend.position = 'bottom',
         legend.title = element_blank(),
         axis.title.x = element_blank(),
         text = element_text(family = 'Century Gothic', color = '#053347'),
         plot.title = element_text(hjust = 0.5, face = 'bold'),
         plot.subtitle = element_text(hjust = 0.5, face = 'bold'),
         plot.caption = element_text(hjust = 0),
         NULL) +
   ylab('') +
   labs(title = 'Daily new cases\' since the 100th case, by country\n',
        subtitle = '7-day moving average',
        caption = '* Only countries with more than 40,000 cases, more than 1,000 cases per million habitants, and with a total\npopulation greater than 5 million people are shown (except for China).\n\nData source: Our World in Data.')

# ggplotly(p,tooltip = c('text','x','text','y')) %>% 
#    layout(legend = list(orientation = 'h', x = 0.1, y = -0.05))
p

```

## Total Deaths by Country

[Text]

[Text]

```{r total_deaths_by_country_p, fig.height=6}

p <- covid1 %>% 
   group_by(continent,.add = TRUE) %>% 
   summarise(total_deaths = max(total_deaths,na.rm = TRUE), .groups = 'keep') %>% 
   mutate('Total deaths' = round(total_deaths/1000,digits = 1)) %>% 
   filter(total_deaths > 1000 & !is.na(continent)) %>% 
   ggplot(aes(x = fct_reorder(location,total_deaths,.desc = TRUE), y = `Total deaths`, fill = continent)) +
   geom_col() +
   scale_y_continuous(labels = scales::comma_format(accuracy = 1)) +
   scale_fill_manual(values = colors_continent) +
   theme_minimal() +
   theme(panel.grid = element_blank(),
         legend.title = element_blank(),
         axis.title.x = element_blank(),
         axis.text.x = element_text(angle = 90,vjust = 0),
         text = element_text(family = 'Century Gothic', color = '#053347'),
         plot.title = element_text(hjust = 0.5, face = 'bold'),
         plot.subtitle = element_text(hjust = 0.5, face = 'bold'),
         plot.caption = element_text(hjust = 0),
         NULL) +
   ylab('Total deaths (\'000)\n') +
   labs(title = 'Total deaths by country\n',
        caption = '* Only countries with more than 1,000 deaths are shown.\n\nData source: Our World in Data.')

ggplotly(p,tooltip = c('text','y'))


```

# Global Mobility

[Text]

[Text]

## Mobility vs. Virus' Spread

[Text]

[Text]

```{r mob_v_covid1, fig.height=6}

covid1 %>% 
   group_by(continent,.add = TRUE) %>% 
   select(continent:new_deaths_per_million, max_cases:max_cases_per_million, population) %>% 
   filter(total_cases >= 100 & population > 1000000) %>% 
   filter(date == min(date) & max_cases_per_million < 10000 & max_cases_per_million > 1000) %>% 
   left_join(select(filter(mobility_raw_data,is.na(sub_region_1)),
                    country_region,date,8:13), 
             by = c('location' = 'country_region', 'date' = 'date')) %>% 
   filter(across(ends_with('baseline'),~!is.na(.))) %>%
   rowwise() %>% mutate(mob_idx = sum(retail_and_recreation_percent_change_from_baseline,
                                      grocery_and_pharmacy_percent_change_from_baseline,
                                      parks_percent_change_from_baseline,
                                      transit_stations_percent_change_from_baseline,
                                      workplaces_percent_change_from_baseline,
                                      na.rm = TRUE)/5) %>% 
   ggplot(aes(x = mob_idx, y = max_cases_per_million, color = continent)) +
   geom_point(aes(size = max_deaths),alpha = 0.25) +
   geom_text(aes(label = location, color = continent), size = 3) +
   scale_size(range = c(.1,20), name = 'Current total deaths', labels = scales::comma) +
   scale_color_manual(values = colors_continent, name = '') +
   scale_y_continuous(labels = scales::comma) +
   theme_minimal() +
   theme() +
   theme(panel.grid = element_blank(),
         text = element_text(family = 'Century Gothic', color = '#053347'),
         plot.title = element_text(hjust = 0.5, face = 'bold'),
         plot.subtitle = element_text(hjust = 0.5, face = 'bold'),
         plot.caption = element_text(hjust = 0),
         NULL) +
   ylab('Current total cases per million habitants\n') + xlab('\n Average decrease in mobility\n\n') +
   labs(title = 'Relation between mobility reduction after\nthe 100th detected case,\nand total cases per million habitants and total deaths\n',
        caption = '*Only countries with total cases per million habitants between 1000 and 10,000, and wit a population\ngreater than 1 million people are shown.\n\nData source: Our World in Data & Google Mobility Reports.')

```

## In-house Activities vs. Virus' Spread

[Text]

[Text]

```{r mob_v_covid2, fig.height=6}

covid1 %>% 
   group_by(continent,.add = TRUE) %>% 
   select(continent:new_deaths_per_million, max_cases:max_cases_per_million, population) %>% 
   filter(total_cases >= 100 & population > 1000000) %>% 
   filter(date == min(date) & max_cases_per_million < 10000 & max_cases_per_million > 1000) %>% 
   left_join(select(filter(mobility_raw_data,is.na(sub_region_1)),
                    country_region,date,8:13), 
             by = c('location' = 'country_region', 'date' = 'date')) %>% 
   filter(across(ends_with('baseline'),~!is.na(.))) %>%
   ggplot(aes(x = residential_percent_change_from_baseline, y = max_cases_per_million, color = continent)) +
   geom_point(aes(size = max_deaths),alpha = 0.25) +
   geom_text(aes(label = location, color = continent), size = 3) +
   scale_size(range = c(.1,20), name = 'Current total deaths', labels = scales::comma) +
   scale_color_manual(values = colors_continent, name = '') +
   scale_y_continuous(labels = scales::comma) +
   theme_minimal() +
   theme() +
   theme(panel.grid = element_blank(),
         text = element_text(family = 'Century Gothic', color = '#053347'),
         plot.title = element_text(hjust = 0.5, face = 'bold'),
         plot.subtitle = element_text(hjust = 0.5, face = 'bold'),
         plot.caption = element_text(hjust = 0),
         NULL) +
   ylab('Current total cases per million habitants\n') + xlab('\n Average decrease in mobility\n\n') +
   labs(title = 'Relation between increment of in-house activities after\nthe 100th detected case,\nand total cases per million habitants and total deaths\n',
        caption = '*Only countries with total cases per million habitants between 1000 and 10,000, and wit a population\ngreater than 1 million people are shown.\n\nData source: Our World in Data & Google Mobility Reports')

```

## Mobility (by type) vs. Virus' Spread {.tabset .tabset-fade .tabset-pills}

[Text]

```{r mob_v_covid3}

plots_lst <- list()

vars <- set_names(names(select(mobility_raw_data,
                               retail_and_recreation_percent_change_from_baseline:residential_percent_change_from_baseline)),
                  c('Retail and Recreation',
                    'Grocery and Pharmacy',
                    'Parks',
                    'Transit Stations',
                    'Work Places',
                    'Residential Activities'))

for (i in names(vars)) {

   plots_lst[[i]] <- covid1 %>% 
      group_by(continent,.add = TRUE) %>% 
      select(continent:new_deaths_per_million, max_cases:max_cases_per_million, population) %>% 
      filter(total_cases >= 100 & population > 10000000) %>% 
      filter(date == min(date) & max_cases_per_million < 10000 & max_cases_per_million > 1000) %>% 
      left_join(select(filter(mobility_raw_data,is.na(sub_region_1)),
                       country_region,date,8:13), 
                by = c('location' = 'country_region', 'date' = 'date')) %>% 
      filter(across(ends_with('baseline'),~!is.na(.))) %>%
      ggplot(aes_string(x = vars[[i]], y = 'max_cases_per_million', color = 'continent')) +
      geom_text(aes(label = location, color = continent, size = max_deaths)) +
      scale_color_manual(values = colors_continent, name = '') +
      scale_y_continuous(labels = scales::comma) +
      scale_size(range = c(1,10), name = 'Current total deaths', labels = scales::comma) +
      theme_minimal() +
      theme() +
      theme(panel.grid = element_blank(),
            text = element_text(family = 'Century Gothic', color = '#053347'),
            plot.title = element_text(hjust = 0.5, face = 'bold'),
            plot.subtitle = element_text(hjust = 0.5, face = 'bold'),
            plot.caption = element_text(hjust = 0),
            NULL) +
      ylab('Current total cases per million habitants\n') + xlab(paste0('\n Average decrease in ',i,' mobility\n\n')) +
      labs(title = paste0('Relation between changes in ', i, ' after\nthe 100th detected case,\nand total cases per million habitants and total deaths\n'),
           caption = '*Only countries with total cases per million habitants between 1000 and 10,000, and wit a population\ngreater than 1 million people are shown.\n\nData source: Our World in Data & Google Mobility Reports.')
   
}

```

```{r, results='asis', fig.height=6}

for (plot in names(plots_lst)) {
   
   cat('\n')

   cat('### ', plot, '\n')

   print(plots_lst[[plot]])

   cat('\n')
}

```

# Acknowledgements

* GitHub Corner: [Tim Holman.](http://tholman.com)

* Icons: [Font Awesome 4.7.0. Created by Font Awesome](https://fontawesome.com/v4.7.0/icons/)
